---
title: "report"
author: "Yuchen Li"
date: "2022-10-20"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Stephen Curry stats in NBA from 2009 to 2021

## Introduction

The dataset I choose from Kaggle is Stephen Curry stats from 2009 to 2022 seasons by MUJIN JO’s newest version. Its contents can be divided into two sections: Golden State Warriors team stats, (such as win/loss, team score, the opponent team name, game date, etc.) and Stephen Curry player stats (such as minutes played in the game, field goals made, three pointers made, rebounds, assists, blocks, etc.). Among all other basketball datasets, I pick this one because of its useability and its focus on my favorite player, Stephen Curry. One thing that makes this dataset useable is its broad variable coverage. This dataset has complete stats dating back to 2009, and the newest data lastly updated in January 23, 2022. Since it contains both overall stats and Stephen Curry specific stats, I can arrive at my conclusions based on analysis from both angles. I would be able to explain problems more comprehensively. This dataset is also meaningful when Golden State Warriors want to change Curry’s contract or decide on if they will trade him to other teams (only hypothetically). 

## Goal

This project focuses on:
1. Stephen Curry's changes in skills over time.
2. Stephen Curry's influence on whole team.
3. Predict Golden State Warriors' future game result based on Stephen Curry's statistics.

## Plots

Below is a bar plot of average minutes Stephen Curry played in each year's pre-season games. We can see that he has lowest on-court time in 2012-2013 pre-season. During this pre-season, Curry just came back from anckle injury rehabilitation. Coach Kerr and the team could want to give him a slow-paced return. I plan to conduct a model of Warrior's game results using Curry's status in the games (ex. minutes played, shooting percentages) as predictors.
```{r echo=FALSE}
readRDS("figures/pre_avg_min_plot.rds")
```

Below is a box plot of Curry's average percent of field goal made in pre-season of each year. There're some interesing pattern:
1. Curry didn't play much in 2013-2013 pre-season, but his percent of field goal made was near average level. It woulde make more sense if compared with team total score and game result. Also, further analysis can be made when taking regular season and post-season records into account.
2. There is a rough trend that Curry's field goal percent in pre-season improved over the years. More analysis can be done to verify this trend and see if other skills (ex. 3-pointer percent, free throw percent) imporved over the years. It would also be meaningful if we take 
```{r echo=FALSE}
readRDS("figures/pre_avg_FGP_plot.rds")

```

## Tables
In progress...

## Dimisionality Reduction
In progress...

## Modeling
In progress...

## Conclusion
In progress...
